---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RHermes

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/RogerGinBer/RHermes/branch/master/graph/badge.svg?token=HL73R4GHFJ)](https://codecov.io/gh/RogerGinBer/RHermes?branch=master)

<!-- badges: end -->

The goal of RHermes is to analyse LC-MS and LC-MS/MS data to identify compounds
in biological or environmental samples. The RHermes workflow works with
both Orbitrap and q-TOF instrument data.


## System requirements

RHermes is quite heavy on the CPU and memory loads of your system, as it has
been developed to use almost all available cores. For that reason you will need:

* At least a 4 core processor
* 16 GB of RAM or more
* An internet connection to perform KEGG queries

## Installation

You can install the released version of RHermes from [Bioconductor](https://bioconductor.org/) with:

``` r
# install.packages("BiocManager")
BiocManager::install("RHermes")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RogerGinBer/Proves")
```

## Setup
RHermes can perform almost all its routines out of the box, but the SOI Blank
Substraction step requires a valid Keras and Tensorflow installation, which rely
on Python.

### Option 1: Default installation
In theory both Keras and Tensorflow can be configured with:

```{r, eval = FALSE}
reticulate::install_miniconda()
keras::install_keras()
tensorflow::install_tensorflow()
```
After which you can check the following:
```{r, eval = FALSE}
tensorflow::tf_config()
model <- keras::load_model_tf(system.file("extdata", "model", package = "RHermes"))
```
If both commands don't yield any error (the "Your CPU supports ..." warning is fine)
you are set to go. If it fails (which can happen in some devices, try Option 2).

### Option 2: Dealing with the Installation yourself
First install Miniconda:
```{r, eval = FALSE}
reticulate::install_miniconda()
```
Now find the Miniconda Prompt in your computer. Instead of relying on the default
r-reticulate environment, type the following to create a new environment:

conda create -n newenv python=3.6 tensorflow keras

When finished, type in R:
```{r, eval = FALSE}
reticulate::use_condaenv("newenv", required = TRUE)
tensorflow::tf_config()
model <- keras::load_model_tf(system.file("extdata", "model", package = "RHermes"))
```

Everything should run smoothly. If not, try manually installing Anaconda from
their website and letting reticulate know where to find the environment.

Also check out Keras and Tensorflow R tutorials.

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(RHermes)
## basic example code
```

