---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RHermes 

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/RogerGinBer/RHermes/branch/master/graph/badge.svg?token=HL73R4GHFJ)](https://codecov.io/gh/RogerGinBer/RHermes?branch=master)

<!-- badges: end -->
<img align = "right" style = "padding-left: 10%; padding-bottom: 10%; padding-top: 2%" 
width = "250px" height = "285px" src = "https://i.postimg.cc/Tw9SvJ11/sticker-No-Bioc.png">

`RHermes` is a broad-scoped targeted metabolomics package designed to analyse 
LC-MS and LC-MS/MS data to identify compounds in biological or environmental
samples. 

The `RHermes` workflow works with both Orbitrap and qTOF instrument data and
comes with an interactive GUI to process and visualize the data.

Whether it's natural products, biomedical or enviormental samples, `RHermes` can
help you improve your matrix characterization. By selecting an appropiate
formula database, you can focus on just the compounds you are interested in and
improve your coverage.

With `RHermes` you can **see** the metabolites in your data and much
more. There's no need to manually calculate m/z's to plot the XIC of different
adducts: with the GUI you are just one click away from a metabolite-centric
plot.

For more info, check out the documentation [here](https://rogerginber.github.io/RHermes/)
and our article [here](https://www.nature.com/articles/s41592-021-01307-z).


## Installation

You can download the development version from
[GitHub](https://github.com/RogerGinBer/RHermes) with:

```{r, eval = FALSE}
if(!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
devtools::install_github("RogerGinBer/RHermes")
```

## Analyzing LC-MS data with RHermes

Once installed, you can use `RHermes` programmatically like this:

```{r example, eval = FALSE}
library(RHermes)
#Generate a Exp object
example <- RHermesExp()

#Set your formula and adduct database
example <- setDB(example, db = "hmdb")

#Process your MS1 files
example <- processMS1(example,
                        system.file("extdata", "MS1TestData.mzML",
                        package = "RHermes"))
#Generate SOIs
example <- findSOI(example, getSOIpar(), 1)

#Generate an IL (Inclusion List)
example <- generateIL(example, 1, ILParam())
```

With the generated inclusion list, you can export it and run a Parallel Reaction
Monitoring (PRM) MS2 experiment to reveal coeluting isomers or use any other MS2 mode
you see fit.

Or start the interactive GUI typing:
```{r, eval = FALSE}
RHermesGUI()
```
In the GUI you will find abundant help pages to guide you along the processing :+1:

Please check the User Guide [vignette](https://rogerginber.github.io/RHermes/articles/RHermes_UserGuide.html) for more detailed info and real examples.

## Database availability

We have compiled some molecular formula open databases ready to be used with RHermes for all sorts of samples: HMDB, ChEBI, NORMAN, LipidMaps, COCONUT, etc. They are freely available at [this Zenodo repository](https://zenodo.org/record/5025560).

## Bug reporting

Suggestions and bug reports are more than welcome at: https://github.com/RogerGinBer/RHermes/issues


## Citation

Please cite this package as:

Giné, R., Capellades, J., Badia, J.M. et al. HERMES: a molecular-formula-oriented method to target the metabolome. Nat Methods 18, 1370–1376 (2021). https://doi.org/10.1038/s41592-021-01307-z



