---
title: "RHermes User Guide"
author: "Roger Giné Bertomeu"
date: "07/08/2020"
output: html_document
vignette: >
  %\VignetteIndexEntry{RHermes User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE)
```

## Intro
RHermes is an untargeted metabolomics data processing package. Its objective is to
identify a customizable set (usually >10K) of metabolites in a biological or
environmental sample.

This document will guide you through the package functions and show you what a typical
RHermes workflow looks like.

## Basic setup
To install RHermes you need an functional R and RStudio installation. 
You can install the package by:
**COMPLETE HERE**


```{r, warning=FALSE}
##Library calls
library(RHermes)
```


## MS1 workflow
### Setting up the RHermesExp object
All the information generated by the package is stored in a single RHermesExp object.
To generate it, you just have to use the S4 constructor:
```{r}
myHermes <- RHermesExp()
```
This object will be the input of most of the package functions and you are expected
to reassign the outputs of the functions to the same RHermesExp object to update it.

Then you can set your MS1 experimental parameters. To do so, we use the function
setExpParam(), which accepts either a custom ExpParam object or a template name:
```{r}
myHermes <- setExpParam(myHermes, params = ExpParam(ppm = 4, res = 120000, ion = "+"))
myHermes <- setExpParam(myHermes, template = "orbi-pos")
```
The following template names are available: "orbi-pos", "orbi-neg", "qtof-pos" and
"qtof-neg". Check ?ExpParam() to see all customizable parameters. All non-specified
ExpParam fields will be left as defaults.

After setting up the ExpParam, we have to decide which set of formulas and adducts 
we want to detect in our data. In general, the less combinations, the faster the 
running time will be. The function used to set the formulas and adducts is setDB():
```{r}
myHermes <- setDB(myHermes, db = "hmdb",) #Loads a small HMDB subset
```
A general set of adducts will be automatically set according to the experimental
parameters. You can choose multicharged and multimer adducts by using the
*adcharge* and *admult* parameters in setDB().

It is also advisable, although not mandatory, to remove unlikely adducts from the
adduct list. To do so, you can check which adducts are selected by just calling
the RHermesExp object and remove adducts with remAd():
```{r, hold = TRUE}
myHermes #Shows a summary of all set parameters and object info
myHermes <- remAd(myHermes, c("M+DMSO+H", "M+IsoProp+H")) #For instance, remove adducts of unused solvents
```


### Processing the MS1 files into PL
The main interface to process files is the function FileProc(), which accepts an
RHermesExp object and a list of paths to the MS1 files to process. All the files 
must share the used experimental parameters (eg. data acquired in different
polarities or in different instruments can't be processed at the same time).


### SOI detection

### SOI filtering/clean-up

### Generating an Inclusion List

## MS2 workflow

## Plotting

## Using the RHermes Shiny app
To start the app, simply type:
```{r, eval=FALSE}
RHermesGUI()
```
And the app should start in your default browser.


