---
title: "1st Andrea Brunner Workflow"
output:
  html_notebook: default
  html_document:
    df_print: paged
    theme: lumen
    toc: true
    toc_float: true
    toc_depth: 4
  pdf_document: default
vignette: >
  %\VignetteIndexEntry{1st Andrea Brunner Workflow}
  %\VignetteEngine{knitr::rmarkdown}
---

# Context

This is the second workflow corresponding to Andrea Brunner's analysis.
We will use the new RHermes S4 classes and methods to illustrate its use.

# MS1 Processing

## Setting the parameters

The first step is to generate an RHermesExp object and set the MS1 experimental 
parameters. To ease the job for the end-user, 4 different templates are
available for both Q-TOF and Orbitrap instruments. The user can also specify 
custom parameters using ExpParam() instead of the template.

New templat

```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "../")  # with something else than `getwd()`
library(RHermes)
require(enviPat)
library(igraph)
require(mzR)
require(magrittr)
require(CHNOSZ)
require(ggplot2)
require(viridis)
require(tidyverse)
library(gtools)
library(RCy3)
library(data.table)
library(keras)
library(BiocParallel)
```

```{r warning=FALSE}
# setwd("../") #For unknitted use of the chunk
myHermes <- RHermesExp()
myHermes <- setExpParam(myHermes,
                        params = ExpParam(ppm = 3.5, res = 120000,
                                          instr = "Orbitrap", minmz = 50,
                                          maxmz = 1200))

negHermes <- RHermesExp()
negHermes <- setExpParam(negHermes,
                         params = ExpParam(ppm = 3.5, res = 120000,
                                          instr = "Orbitrap", minmz = 50,
                                          maxmz = 1200, ion = "-"))
```

Processing the RHermesExp object will set timestamps of the performed 
operations. You can use readTime(object) to see them.
```{r, message = 1}
readTime(myHermes)
```

Now we will set the formulas and adducts to consider in our experiment. The
default database is the Human Metabolome Database (HMDB), but NORMAN is also 
available and is particularly useful for environmental pollutant profiling.

The IDs to use for the two databases are, respectively, "hmdb", and "norman".

RHermes directly checks the polarity used in the experiment (set in the 
previous step) and selects the corresponding adducts.

The user can also select the characteristics of the adducts considered, in 
particular one can select the maximum charge (+, 2+, etc.) and the maximume
multiplicity of the adducts (M, 2M, 3M, etc.). The defaults are 1 for both
charge and multiplicity.

```{r, message=seq(2,3), warning= FALSE}
myHermes <- setDB(myHermes, db = "custom", filename = "D:/norman.xls")
myHermes <- remAd(myHermes, myHermes@metadata@ExpParam@adlist$adduct[-c(1:5)])

negHermes <- setDB(negHermes, "custom", filename = "D:/norman.xls")
negHermes <- remAd(negHermes, negHermes@metadata@ExpParam@adlist$adduct[-c(1,7)])
readTime(myHermes)
```

## File processing into PLs

Once the basic parameters have been set it is time to process some files. To do
this we just need to use the function processMS1 with the names of the files we 
want processed. There is the possibility to process labelled data so that all 
possible C isotopes are searched but, in that case, the labelled files must be 
processed in a separate function call (or just process everything in _labelled 
mode_, just bear in mind it is significantly slower).

```{r, message = 1, warning = FALSE}
dir <- "D:/ABrunner"
fil <- list.files(dir, pattern = ".*pos.*.mzML", full.names = TRUE)[c(1,4,7)]
myHermes <- processMS1(myHermes, files = fil)

fil <- list.files(dir, pattern = ".*neg.*.mzML", full.names = TRUE)[c(1,4,7)]
negHermes <- processMS1(negHermes, files = fil)
```
## SOI generation

Once every file has been processed and all signals that match your F/A
combinations have been registered into _Peak Lists_ the next step will be SOI
detection.

As described in the paper, RHermes detects SOI (_Scans of Interest_) by applying
one or multiple filters to the data points registered in a PL. To provide an 
almost hands-off experience, we have designed multiple filter templates that can
be easily accessed and customized. They are:

* _single_ and _single-x_
* _double_ and _double-x_
* _triple_ and _triple-x_

Where the "-x" suffix means "extended" and is to be used when dealing with long
chromatographic runs where compounds can elute during a long time. Double and
triple filter modes use staggered, partially-overlapping time segments to avoid
signal splitting.

In case you need more filters or if you just want to customize them you can find
the "SOITemplates.csv" file at /app/www in the RHermes package folder. Just add
more lines with your desired parameters using the same ID in the first column
and it will be available to getSOIpar().

It is likely that you may want to test some filtering parameters before
commiting to perform a set of MSMS runs. With this in mind, findSOI() allows
the input of multiple parameter objects in the form of a list (or c()). One 
 thing to keep in mind is that, in the case the length of parameters is
shorter than the number of SOI lists to generate, **only** the last parameter of
the list will be recycled.

Another feature that RHermes offers is the possibility to select a PL as
blank and _substract_ those signals when processing another PL. This step is 
optional but really advisable, since it will reduce the number of MSMS 
injections needed afterwards. It works like this:

 * You can input a numeric vector telling the program which PLs you want to use
 as blank. 
 * If you don't enter anything it doesn't perform any substraction by
 default, so don't worry.
 * 0 means _No blank substraction_, and it is important because both the 
 _fileID_ and _against_ vectors will be matched **one-to-one**.
 


```{r , warning=FALSE}
setwd("../") #For unknitted use of the chunk
s <- getSOIpar("double")
myHermes <- findSOI(myHermes, params = s, fileID = c(1,2,2,3,3,3),
                      against = c(0,0,1,0,1,2))
saveRDS(myHermes, file = "D:/HermesResults/SurfaceWater/PosSOI.rds")

for(i in 1:6){
  myHermes <- cleanSOI(myHermes, soiid = i, minint = 10000, TRUE)
  myHermes <- removeISF(myHermes, id = i)
}

saveRDS(myHermes, file = "D:/HermesResults/SurfaceWater/PosSOIFilter.rds")
```

```{r, warning=FALSE}
s <- getSOIpar("double")
negHermes <- findSOI(negHermes, params = s, fileID = c(1,2,2,3,3,3),
                      against = c(0,0,1,0,1,2))
saveRDS(negHermes, file = "D:/HermesResults/SurfaceWater/NegSOI.rds")

for(i in 1:6){
  negHermes <- cleanSOI(negHermes, soiid = i, minint = 10000, TRUE)
  negHermes <- removeISF(negHermes, id = i)
}
saveRDS(negHermes, file = "D:/HermesResults/SurfaceWater/NegSOIFilter.rds")
```

## Quality control - Plotting
```{r}
```


## Generating an Inclusion List
Once you are satisfied with a SOI list it is time to generate an inclusion list
to perform MSMS fragmentation of the SOI and proceed with the workflow.

Select the desired SOI list and the IL parameters:

 * **mode**: We offer three different inclusion list modes which offer some 
flexibility when designing the number of MSMS injections. "Full" will use all
SOIs, "only" will just pick the SOIs that are associated with a list of adducts
ad and "priorizate" will, if possible, only pick the first adduct in ad for a
certain formula, then the next, and so on. This last mode avoids fragmenting 
redundant adduct signals.
 * **ad**: The list of adducts to pick/prioritize. Check your 
object@metadata@ExpParam@adlist to ensure proper spelling.
 * **filtermz**: **Really** important, is the mz tolerance of your quadrupole filter.
 It has to be properly set so that SOIs overlapping in a mz-rt neighborhood are 
 grouped together into a single IL entry. Feel free to experiment, but larger mz
 values can yield convoluted the MSMS spectra. 

```{r}
myHermes <- generateIL(myHermes, id = 3, par = ILParam(priorization = "full"))
myHermes <- generateIL(myHermes, id = 3, par = ILParam(priorization = "yes", ad = c("M+H", "M+NH4", "M+CH3OH+H")))
myHermes <- generateIL(myHermes, id = 6, par = ILParam(priorization = "full"))
myHermes <- generateIL(myHermes, id = 6, par = ILParam(priorization = "yes", ad = c("M+H", "M+NH4", "M+CH3OH+H")))
myHermes <- generateIL(myHermes, id = 4, par = ILParam(priorization = "full"))
myHermes <- generateIL(myHermes, id = 4, par = ILParam(priorization = "yes", ad = c("M+H", "M+NH4", "M+CH3OH+H")))
saveRDS(myHermes, file = "D:/HermesResults/SurfaceWater/PosSOIFilter.rds")


negHermes <- generateIL(negHermes, id = 3, par = ILParam(priorization = "full"))
negHermes <- generateIL(negHermes, id = 3, par = ILParam(priorization = "yes", ad = c("M-H")))
negHermes <- generateIL(negHermes, id = 6, par = ILParam(priorization = "full"))
negHermes <- generateIL(negHermes, id = 6, par = ILParam(priorization = "yes", ad = c("M-H")))
negHermes <- generateIL(negHermes, id = 4, par = ILParam(priorization = "full"))
negHermes <- generateIL(negHermes, id = 4, par = ILParam(priorization = "yes", ad = c("M-H")))
saveRDS(negHermes, file = "D:/HermesResults/SurfaceWater/NegSOIFilter.rds")
```

Creating an inclusion list will generate an instance of an MS2Exp, a container
which will serve to store your IL, the MSMS data you will later acquire and the
identifications that arise from said data.

With the inclusion list ready you can just export it to a csv format that can be
used as input for the MSMS experiment. Just select the IL to export, where to 
save it, the maximum number of ions to monitor at the same time and whether
you want to get separate csv files or just one csv.

```{r}
# exportIL(myHermes, 1, folder = getwd(), maxOver = 10, sepFiles = TRUE)
# exportIL(negHermes, 1, folder = getwd(), maxOver = 10, sepFiles = TRUE)
```

# MS2 Processing

After performing the MSMS runs with the exported IL you can process the
generated data with RHermes and obtain compound identifications.

## Using processMS2

```{r eval=FALSE}
ms2f <- list.files("D:/ABrunner/MSMSdata", pattern = ".*pos.*.mzML", full.names = TRUE)[seq(1,31)]
myHermes <- processMS2(myHermes, 1, MS2files = ms2f, sstype = "regular", referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)
myHermes <- processMS2(myHermes, 2, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)


ms2f <- list.files("D:/ABrunner/MSMSdata", pattern = ".*pos.*.mzML", full.names = TRUE)[seq(33,45)]
myHermes <- processMS2(myHermes, 3, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)
myHermes <- processMS2(myHermes, 4, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)

ms2f <- list.files("D:/ABrunner/MSMSdata", pattern = ".*pos.*.mzML", full.names = TRUE)[c(1:31, seq(33,45))]
myHermes <- processMS2(myHermes, 5, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)
myHermes <- processMS2(myHermes, 6, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)

saveRDS(myHermes, file = "D:/HermesResults/SurfaceWater/MS1_pos_Ident_Water.rds")

```



```{r eval=FALSE}
ms2f <- list.files("D:/ABrunner/MSMSdata", pattern = ".*neg.*.mzML", full.names = TRUE)[seq(3,33)]
negHermes <- processMS2(negHermes, 1, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)
negHermes <- processMS2(negHermes, 2, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)


ms2f <- list.files("D:/ABrunner/MSMSdata", pattern = ".*neg.*.mzML", full.names = TRUE)[seq(35,45)]
negHermes <- processMS2(negHermes, 3, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)
negHermes <- processMS2(negHermes, 4, MS2files = ms2f,sstype = "regular",  referenceDB = "D:/MS2ID_B2R_20201113_083214.rds", useDB = TRUE, mincos = 0.5)
saveRDS(negHermes, file = "D:/HermesResults/SurfaceWater/MS1_neg_Ident_Water.rds")
```

```{r}
sink("D:/HermesResults/SurfaceWater/sessionInfo.txt")
sessionInfo()
paste("Date: ", date())
sink()
rm(list = ls())
```



